My title

### 2024-03-11 paper/paper.qmd
model and data so far

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
  - Aamishi Sundeep
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
date: today
date-format: long
abstract: "The 2024 U.S. Presidential Election ..."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(janitor)
library(rstanarm)
library(readr)
library(ggplot2)
```

@@

Delegates to secure the nomination, Nikki Haley has dropped out of GOP

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

@@

Our data is from ...

@@

@@

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

@@

## First discussion point {#sec-first-point}

@@

## Second discussion point

## Third discussion point

@@

## Posterior predictive check

@@

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
  - Aamishi Sundeep Avarsekar
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
date: today
date-format: long
abstract: "The 2024 U.S. Presidential Election ..."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(janitor)
library(rstanarm)
library(readr)
library(ggplot2)
library(knitr)
library(arrow)
```

@@

Delegates to secure the nomination, Nikki Haley has dropped out of GOPYou can and should cross-reference sections and sub-sections. We use @citeR and

@@

```{r}
library(arrow)
library(dplyr)
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
# poststrat_analysis_data <- read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
print(survey_analysis_data)

proportion_biden <- survey_analysis_data %>%
  filter(vote_biden == 1) %>%
  summarise(proportion = n() / nrow(survey_analysis_data))

proportion_biden_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(proportion_biden = mean(vote_biden == 1, na.rm = TRUE))

proportion_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(proportion = n() / nrow(survey_analysis_data))

proportion_biden_by_state
proportion_by_state

# only includes 7 observations from hawaii
# Hawaii is historically a Democratic stronghold
```

```{r}
#| label: fig-survey_presidential_preferences
#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
#| echo: false
survey_analysis_data |>
  ggplot(aes(x = education_level, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-survey_presidential_preferences1
#| fig-cap: The Distribution of Presidential Preferences, by gender and race
#| echo: false

survey_analysis_data |>
  ggplot(aes(x = races, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Race",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@@

@citeIPUMS

@@

The logistic regression used to predict ... is as follows:

```{=tex}
\begin{align}
y_i &= \beta_0 + \beta_1 \times x_{sex} + \beta_2 \times x_{age\_bracket} + \beta_3 \times x_{races} + \beta_3 \times x_{education\_level} + \beta_4 \times x_{state} + \beta_5 
\times x_{urban} + \beta_5 \times x_{race\_black}
\end{align}
```
We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

We use the `predict()` function in R @citeR to apply our logistic regression model to the ACS Census data @citeIPUMS.

Model weaknesses: logistic regression output is binary so it does not include the options of not voting or voting for a third-party candidate.

@@

presidential_vote_model <-
  readRDS(file = here::here("models/presidential_vote_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of presidential vote based on ..."
#| warning: false

modelsummary::modelsummary(
  list(
    "Presidential Vote Model" = presidential_vote_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

## Popular Vote Projection {#sec-first-point}

@@

## Electoral College Projection

## Polarization and America's Urban-Rural Divide

@@

## Correlation Map

@@

Model plot (still need to fix this)

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelplot1
#| tbl-cap: "Model Plot"
#| warning: false

# modelplot(presidential_vote_model, conf_level = 0.9) +
#   labs(x = "90 per cent credibility interval")
